---
title: "Influence of transmission type on MPG"
output: pdf_document
geometry: margin=1.5cm
---
Executive summary
-------------
There is a statistically significant difference between mpg values for different types of transmission. Analysis based on linear models shows that automatic transmission gives 2.9358 more miles per gallon comparing to manual transmission.

Exploratory analysis
------------------------
Loading data
```{r}
library(datasets)
data(mtcars)

mtcars$am<-factor(mtcars$am)
```
We see on graph (Appendix 1) considerable difference among mpg means for two types of transmission.
Let's see if this difference is statisitically significant.
```{r}
fit<-lm(mpg~am, mtcars)
summary(fit)
```
Intercept here represents manual transmission, while automatic gives significant (at p-value<0.05) increase.
But this model explains only 36% of variance, besides chances are the coefficient we get is not true because of the hidden correlated variables. So we gonna add some predictors to the model.

Regression analysis
-------------------
Let's begin adding all the variables as regressors.
```{r}
fit<-lm(mpg~., mtcars)
summary(fit)$coef
```
Now we'll be removing predictors one by one according to the highest p-values until we get meaningful model.
```{r}
fit<-lm(mpg~.-cyl, mtcars)
fit<-lm(mpg~.-cyl-vs, mtcars)
fit<-lm(mpg~.-cyl-vs-gear, mtcars)
fit<-lm(mpg~.-cyl-vs-gear-carb, mtcars)
fit<-lm(mpg~.-cyl-vs-gear-carb-drat, mtcars)
fit<-lm(mpg~.-cyl-vs-gear-carb-drat-disp, mtcars)
fit<-lm(mpg~.-cyl-vs-gear-carb-drat-disp-hp, mtcars)
summary(fit)
```
Now we have model with wt, qsec and am as regressors, wich explains about 85% of variance.
We see that automatic transmission gives now 2.9358 increase in mpg with other predictors keeping fixed.

Residuals and outliers diagnostics
---------------------------
On the residuals plot (Appendix 2) we see no pattern, residuals are approximately identically distributed above and below 0, which means our model is quite good.
On the leverage vs residuals plot (Appendix 3) we see no signifaicant leverage for any point in our data, which means outliers don't have much influence on our model.

Appendix 1
--------------
```{r}
library(ggplot2)
g<-ggplot(mtcars, aes(x=factor(am), y=mpg, fill=factor(am)))+
geom_violin()+
xlab("Type of transmission")+ylab("MPG")+
geom_abline(intercept=mean(mtcars$mpg[mtcars$am==0]), slope=0, size=2)+
geom_abline(intercept=mean(mtcars$mpg[mtcars$am==1]), slope=0, size=2)
g
```
\newpage 

Appendix 2
--------------
```{r}
plot(fit,which=1)
```
\newpage

Appendix 3
--------------
```{r}
plot(fit,which=5)
```